---
title: "stat_project"
output: html_document
---

```{r}
df = read.csv("car_price_dataset_clean_stat.csv")
df
```

```{r}
2
df$Car_ID <- NULL

# Identify numeric columns for scaling
num_cols <- c("Model_Year", "Engine_Size", "Mileage", "Doors", "Owner_Count", "Horsepower")

# Apply scaling (Standardization)

df[num_cols] <- lapply(df[num_cols], scale)

# Categorical Variables (Convert to factors)
df$Brand <- as.factor(df$Brand)
df$Fuel_Type <- as.factor(df$Fuel_Type)
df$Transmission <- as.factor(df$Transmission)
df
```
```{r}
#split
dim(df)
set.seed(123) 
sample_idx <- sample(1:nrow(df), 0.8 * nrow(df))
train_data <- df[sample_idx, ]
test_data  <- df[-sample_idx, ]
dim(train_data)
dim(test_data)
```
```{r}
library(corrplot)
cor_matrix <- cor(train_data[num_cols], method="spearman")

corrplot(cor_matrix, 
         method = "color",       
         type = "upper",         
         addCoef.col = "black", 
         tl.col = "black",      
         tl.srt = 45,           
         diag = FALSE           
)

```
```{r}
cor_matrix_2 <- cor(train_data[num_cols], method="pearson")

corrplot(cor_matrix_2, 
         method = "color",       
         type = "upper",         
         addCoef.col = "black", 
         tl.col = "black",      
         tl.srt = 45,           
         diag = FALSE           
)

```
There is no strong linear relationship between these predictors
```{r}
# first model takes only numeric features as predictors 
base_mod1 <- lm(Price ~ Mileage + Model_Year + Horsepower + Engine_Size + Owner_Count + Doors, data = train_data)
summary(base_mod1)
plot(base_mod1)

# second model takes all features as predictors 
base_mod2 <- lm(Price ~ Mileage + Model_Year + Horsepower + Engine_Size + Owner_Count + Doors + Fuel_Type + Transmission + Brand , data = train_data)
```
```{r}
summary(base_mod2)
plot(base_mod2)
```
```{r}
install.packages("car")
library(car)
```


```{r}
# check Vif
vif_base1= vif(base_mod1)
print(vif_base1)
```
```{r}
vif_base2= vif(base_mod2)
print(vif_base2)
```

All values are besicly 1 so there is no multicollinearity between predictors
```{r}
# for first enhanced model I will use only features with ***  
model_enhanced_1 <- lm(Price ~ Mileage + Model_Year + Horsepower + Engine_Size + Owner_Count,  data = train_data)
plot(model_enhanced_1)
summary(model_enhanced_1)

```
```{r}
vif_enhanced_1= vif(model_enhanced_1)
print(vif_enhanced_1)
```
Also there is no multicolinearity 

```{r}
crPlots(model_enhanced_1)
```
Analysis of the Component + Residual Plots confirms that the primary predictors—Mileage, Horsepower, Model_Year,Owner_count and Engine_Size—exhibit strong linear relationships with the target variable. The smoothed loess lines (pink) closely follow the linear null hypothesis lines (blue), indicating that polynomial transformations are not required for these features."

```{r}
# for the second enhenced model i will use interaction terms
model_enhanced_2 <- lm(Price ~ Mileage + Model_Year + Horsepower + Engine_Size + Owner_Count + Mileage * Model_Year + Horsepower * Model_Year + Horsepower * Model_Year +Mileage * Owner_Count,data = train_data)
summary(model_enhanced_2)
```


```{r}
summary(model_enhanced_2)
```

```{r}
AIC(base_mod1,base_mod2,model_enhanced_1,model_enhanced_2)
```


```{r}
predictions <- predict(model_enhanced_1, newdata = test_data)


rmse_test <- sqrt(mean((results$Actual - results$Predicted)^2))

mae_test <- mean(abs(results$Actual - results$Predicted))

plot(test_data$Price, predictions,
     xlab = "Actual Price (Truth)",
     ylab = "Predicted Price (Model)",
     main = "Test Set: Actual vs. Predicted",
     pch = 16, col = rgb(0, 0, 1, 0.5))

abline(0, 1, col = "red", lwd = 2)

```

